---
title: "Bigquery"
output:
  pdf_document: default
  html_document:
    df_print: paged
---


```{r}
library(bigrquery)
library(tidyverse)
```

When using bigrquery interactively, you’ll be prompted to authorize bigrquery in the browser. 

- login the Google Cloud Platform
- create a new project
- enable BigQuery API
- add public data


```{r, eval = FALSE}
bigrquery::bq_auth().
```

```{r, cache = TRUE}
# replace it with your project id
project <- "adept-vigil-269305"
result <- bq_project_query(
  project,
  "SELECT * FROM `bigquery-public-data.samples.gsod` LIMIT 100;")
```

```{r, cache = TRUE}
bq_table_download(result)
```


# Upload dataset

You could upload via the web interface or using `bq_` functions.


```{r, eval = FALSE}
mydataset <- bq_dataset(project, "mydataset")
bq_dataset_create(mydataset)
bq_dataset_exists(mydataset)
```

Let's try to upload the `mtcars` dataset and pretend that it is huge.


```{r, eval = FALSE}
ta <- bq_table(mydataset, "mtcars")
bq_table_create(
  ta,
  friendly_name = "Motor Trend Car Road Tests",
  description = "The data was extracted from the 1974 Motor Trend US magazine",
  labels = list(category = "example")
)
bq_table_exists(ta)

cars <- mtcars %>% 
  mutate(cyl = as_factor(cyl), vs = as_factor(vs), am = as_factor(am))
bq_table_upload(ta, cars, fields = as_bq_fields(cars))
```

Now, let's have some fun.

There are three interfaces provided by `bigrquery`.
- Low level API overy REST
- DBI
- dplyr


*bq_*
```{r}
result <- bq_project_query(
  project,
  "SELECT * FROM `adept-vigil-269305.mydataset.mtcars` where `mpg` < 30")
bq_table_download(result)
```

```{r}
library(DBI)
con <- dbConnect(
  bigquery(),
  project = project,
  dataset = "mydataset"
)
```

*DBI*
```{r}
con %>% dbGetQuery("SELECT * FROM `adept-vigil-269305.mydataset.mtcars` WHERE `mpg` < 30")
```

*dplyr*
```{r}
con %>% tbl("mtcars") %>% 
  filter(mpg < 30) %>% 
  collect()
```


```{sql connection = con}
SELECT * FROM `adept-vigil-269305.mydataset.mtcars` WHERE `mpg` < 30;
```


# Running linear regression in Bigquery

```{sql connection = con, eval = FALSE}
CREATE TABLE `mydataset.mtcars2` AS
  SELECT *,
  RAND() as `train`
  FROM `adept-vigil-269305.mydataset.mtcars`
```



```{sql connection = con, eval = FALSE}
CREATE MODEL `mydataset.mtcars_model`
OPTIONS
  (model_type='linear_reg',
    input_label_cols=['mpg']) AS
SELECT
  `mpg`,
  `cyl`,
  `disp`,
  `hp`,
  CAST(`gear` AS string) AS `gear`
FROM
  `adept-vigil-269305.mydataset.mtcars2`
WHERE
  `train` < 0.9  -- select rows randomly
```

If you want to delete the model
```{sql connection = con, eval = FALSE}
DROP MODEL `mydataset.mtcars_model`;
```

To do prediction

```{sql connection = con, eval = FALSE}
SELECT * FROM ML.PREDICT(MODEL `adept-vigil-269305.mydataset.mtcars_model`, (
  SELECT
    `cyl`,
    `disp`,
    `hp`,
    CAST(`gear` AS string) AS `gear`
  FROM `adept-vigil-269305.mydataset.mtcars2` WHERE `train` >= 0.9
))
```


# Reference

BigQuery: https://cloud.google.com/bigquery-ml/docs/reference/standard-sql
