---
title: "A crash course on Rcpp"
output:
  pdf_document: default
  html_document: default
date: "01-28-2020"
---


# System requirements

To run c++ code in R, you will need to first install a workable toolchain.

- Windows
  - Install [Rtools](https://cran.r-project.org/bin/windows/Rtools/)
- macOS
  - Run `which clang` in terminal to see if you have already installed it
  - If not, run `xcode-select --install` to install it.
- Linux
  - It depends

In R, we will need to install the package `Rcpp`.


The following sections will teach you the basics by translating simple R functions to their C++ equivalents

- No inputs, scalar output
- Scalar input and scalar output
- Vector input and scalar output
- Vector input and vector output
- Matrix input and vector output


## No inputs, scalar output

This function has no arguments and always returns the integer 1:
```{r}
oneR <- function() 1L
```

```{Rcpp}
// [[Rcpp::export]]
int oneC() {
  return 1;
}
```

- You must declare the type of output the function returns. 
- Every statement is terminated by a `;`
- R doesn't have the idea of scalar variables but scalars and vectors in C++ are different. 

| R type | Rcpp type|
| -------| ------|
| int scalar | int |
| double scalar | double |
| logicial scalar | bool |
| character scalar | String |
| int vector | IntegerVector |
| double vector | NumericVector |
| logicial vector | LogicalVector |
| character vector | CharacterVector |
  

```{r}
oneR()
oneC()
```

## Scalar input, scalar output

The next example function implements a scalar version of the sign() function which returns 1 if the input is positive, and -1 if it’s negative:

```{r}
signR <- function(x) {
  if (x > 0) {
    1
  } else if (x == 0) {
    0
  } else {
    -1
  }
}
```

```{Rcpp}
// [[Rcpp::export]]
int signC(int x) {
  if (x > 0) {
    return 1;
  } else if (x == 0) {
    return 0;
  } else {
    return -1;
  }
}
```

- We declare the type of each input in the same way we declare the type of the output. 
- The `if` syntax is identical. 
- The `while` loop syntax is almost the same, we use `break` to break the loop but use `continue` instead of `next` skip current iteration.

```{Rcpp}
// [[Rcpp::export]]
int while_ex(int x) {
  while(x < 9) {
    x = x + 1;
    if (x == 7) {
      break;
    }
  }
  return x;
}
```

```{r}
while_ex(1)
while_ex(7)
while_ex(9)
```

## Vector input, scalar output

It is where things get compicated.

```{r}
# it is the inefficient version of the function `sum`
sumR <- function(x) {
  total <- 0
  for (i in seq_along(x)) {
    total <- total + x[i]
  }
  total
}
```

```{Rcpp}
// The first two lines import the namespace Rcpp for things such as `NumericVector`
#include <Rcpp.h>
using namespace Rcpp;

// [[Rcpp::export]]
double sumC(NumericVector x) {
  int n = x.size();
  double total = 0;
  for(int i = 0; i < n; i++) {
    total += x[i];
  }
  return total;
}
```

- `NumericVector` is the Rcpp type of a double vector. To find its length, we could use the `.size()` method of `NumericVector`.
- We need to specify the type of each variable. `int n` and `double total`.
- The `for` loop starts with `i = 0` and ends if `i < n`. 
- After each iteration, we increment the value of `i` by one, `i++` increases the value of `i` by 1
- In C++, VECTOR INDICES START AT 0!!! The first element is `x[0]` and the last element is `x[n - 1]`;
- Use `=` for assignment, not `<-`
- `total += x[i]` means `total = total + x[i]`. Similarly, there are `-=`, `*=`, and `/=`.


Let's benchmark the performances
```{r, cached = TRUE}
x <- runif(1e3)
result <- bench::mark(
  sum(x),
  sumC(x),
  sumR(x)
)
ggplot2::autoplot(result)
```
Of course `sum` is the fastest because it is highly optimized.



## Vector input, vector output

This is a function that computes the Euclidean distance between a value and a vector of values:

```{r}
pdistR <- function(x, ys) {
  sqrt((x - ys) ^ 2)
}
x <- runif(1e5)
pdistR(0.5, x)
```

```{Rcpp}
#include <Rcpp.h>
using namespace Rcpp;

// [[Rcpp::export]]
NumericVector pdistC(double x, NumericVector ys) {
  int n = ys.size();
  NumericVector out(n);

  for(int i = 0; i < n; ++i) {
    out[i] = sqrt(pow(ys[i] - x, 2.0));
  }
  return out;
}
```

- We create a new numeric vector of length `n` with a constructor: `NumericVector out(n);`. 
- Another useful way of making a vector is to copy an existing one: `NumericVector zs = clone(ys);`.
- C++ uses `pow()`, not `^`, for exponentiation.
- C++ is faster because it avoids the creation of temperary object like `x - ys` and `(x - ys)^2`.

```{r}
y <- runif(1e6)
bench::mark(
  pdistR(0.5, y),
  pdistC(0.5, y)
)
```

We saved about `5ms` by rewriting the R function in C++. Assume it takes 5 mins for you to write that function, you’d need to run it ~60 000 times to make rewriting worthwhile. 


# Reference

- Rcpp chapter of Advanced R https://adv-r.hadley.nz/rcpp.html
- Rcpp Quick Reference Guide http://dirk.eddelbuettel.com/code/rcpp/Rcpp-quickref.pdf
