---
title: "Shell tutorial"
output:
  pdf_document: default
  html_document: default
date: "02-06-2020"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

A Shell is a program that enables user to interact with the operating system. There are a few common implementaion of a shell.

- bash
  - the most common shell
  - default in linux
  - used to be the default shell on macOS
- zsh
  - the default shell in macOS 10.15 Catalina
- powershell
  - primarily used in windows
  
We will focus on `bash` (and `zsh` is `bash` compatible).


There are a few program to launch the shell.

- linux
  - depends on the desktop environment, it could be gnome terminal, konsole etc..
- macOS
  - the default Terminal
  - iTerm2
- Windows
  - Powershell
  - Cmd (which runs DOS)
  - Cmder
- RStudio terminal
- Rmarkdown (for bash script)


## Commands to cover

### File navigation

- ls
  
  A usful alias is
  ```
  alias ls='ls --color=auto -F'
  ```
  
- pwd
- cd


### File manipulation

- touch
- mkdir
- mv
- cp
- rm

### Utility

- echo
- cat
- which
- export


## SSH

> The SSH protocol (also referred to as Secure Shell) is a method for secure remote login from one computer to another.

To use ssh, we need to install ssh.

- linux and macOS
  - you should have a the ssh program already
- Windows
  - if you have install git for windows, we should have git-bash and ssh installed.
  

### Login to your peloton account

```
ssh sta141c-NUMBER@peloton.cse.ucdavis.edu
```

Unfortunately, you will need to copy and paste password everytime you login. (I will teach you a way to do it in R without typing password each time).

  
### File editing over ssh


Editing a file over SSH is not an easy task. 

- use a text editor in the system
- edit it locally and upload it to the server
  - either use `scp` functions 
  - or a software, for example "Cyber Duck" for macOS and WinScp for Windows.

### `scp` function to copy files from to to server

To copy a local file `file.txt` to remote `~/foo/file.txt`
```
scp file.txt sta141c-98@peloton.cse.ucdavis.edu:~/foo/
```

To copy a local directory to remote `~/foo/local_directory`
```
scp -r local_directory sta141c-98@peloton.cse.ucdavis.edu:~/foo/
```

To copy a local file `file.txt` from remote `~/foo/file.txt`
```
scp sta141c-98@peloton.cse.ucdavis.edu:~/foo/file.txt ./
```

To copy a directory from remote `~/foo/`
```
scp -r sta141c-98@peloton.cse.ucdavis.edu:~/foo/ ./
```


### The package `ssh` and `keyring`.

```{r}
library(keyring)
# only need to do it once
key_set("PELOTONPASSWORD")
```

```{r}
library(ssh)
session <- ssh_connect(
  "sta141c-98@peloton.cse.ucdavis.edu", 
  passwd = keyring::key_get("PELOTONPASSWORD"))
session
```
```{r}
session %>% ssh_exec_wait(
  "ls -A"
)
```


```{r}
session %>% ssh_exec_wait("mkdir -p shell")
session %>% scp_upload("shell.Rmd", to = "shell/")
session %>% scp_download("shell/shell.Rmd", to = "foo/")
```

```{r}
ssh_disconnect(session)
```


# Reference

- Episode 1 to 3 of http://swcarpentry.github.io/shell-novice/
